An Analysis of Health & Lifestyle Factors Associated with Sleep Quality

### 2024-05-03 products/manuscript/Manuscript.qmd
5/3- check

For instance, in Sun et al. (2015), a sample of Chinese individuals
showed a correlation between obesity and worsened sleep quality in men
but not women [@sun_poor_2016]. We are interested in seeing if this 
trend in present in
our dataset; however, given that most of the observations are of men, we
expect to see a correlation between lower sleep scores and obesity in
our data. Furthermore, it has been shown that age and gender have a
notable effect on sleep quality (Madrid-Valero et al. 2016)[@madrid-valero_age_
2017]. Women seem to experience a deterioration in sleep quality as they age; 
however,while this trend is still present in men, it is notably less consistent
and can vary dramatically between individuals. We expect to see similar
trends in our dataset, but as stated previously, our observations
contain significantly more men than women. Finally, we wish to test
whether or not certain occupational industries have different effects on
sleep quality. It has been shown that managerial positions tend to have
the worst sleep quality among civilian sector workers while 24 jobs that
have rotating shifts tend to have the worst sleep quality [@luckhaupt_prevalence
_2010]. Our dataset contains observations of mostly
white-collar/service-based jobs, so we expect to find similar trends
upon analysis.

Additionally, based on the findings of Sun et. al that obesity can
negatively impact sleep quality, we also expect that higher levels of
blood pressure will result in poorer sleep quality [@sun_poor_2016]. 
The American SleepFoundation states that both short and long sleep duration 
increase risk of hypertension in adults [@liu_association_2022]. This draws 
interesting focus to the circular nature of sleep quality and hypertension. 
Those with hypertension also face a higher risk of insomnia, a disorder
affecting the ability to fall asleep, maintain sleep, and sleep without
continuous disturbance [@liu_association_2022]. Short sleep duration is also
positively associated with stress level [@noauthor_stress_nodate]. Survey findings
indicate that the average American sleeps 6.7 hours per night, which is
well below the minimum recommended duration seven to nine hours (APA,
2013). Similar to the relationship between hypertension and sleep
quality, as stress level increases, sleep duration decreases thereby
causing poorer sleep quality. Poorer sleep quality then leads to higher
stress, greater risk of hypertension, and shorter sleep duration.

@@

The dataset originally contained 373 observations for 13 variables
covering a broad spectrum of health and lifestyle variables associated
with sleep quality. The curators of the dataset did not provide
information on how or where this data was collected, so we solely relied
on the codebook for variable definitions. We have added this codebook to
the supplementary materials file for ease of access but you can also
find it on the Kaggle page for this dataset. There were no missing
values or erroneous variables in the dataset that we needed to remove.
The "Blood.Pressure" variable was transformed to reflect the individual
systolic and diastolic blood pressure measurements from each subject. We
then chose to add an additional variable, labeled 'cat_bp' to reflect
the categorical blood pressure status of a subject based on the
[American Heart Association
guidelines](https://www.heart.org/en/health-topics/high-blood-pressure) [@noauthor_understanding_nodate].
Additionally we created a variable named 'StepsGroup' to determine the
categorical activity level of subjects in the dataset based on these
guidelines provided by [10,000
steps](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps
/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,
day%205%20Highly%20active%20is%20more%20than%2012%2C500) [@noauthor_counting_
nodate].
Finally, we created an additional categorical variable called
'PhysicalActivityGroup' to categorically represent the level of physical
activity reported by a subject. The levels are differentiated by 30
minute intervals of weekly physical activity. The outcome of interest is
noted as "Quality.of.Sleep" that we will reference as Sleep Quality or
Quality of Sleep throughout this report.

>>>>

For instance, in Sun et al. (2015), a sample of Chinese individuals
showed a correlation between obesity and worsened sleep quality in men
but not women [@sun_poor_2016]. We are interested in seeing if this 
trend in present in
our dataset; however, given that most of the observations are of men, we
expect to see a correlation between lower sleep scores and obesity in
our data. Furthermore, it has been shown that age and gender have a
notable effect on sleep quality[@madrid-valero_age_2017]. 
Women seem to experience a deterioration in sleep quality as they age; 
however,while this trend is still present in men, it is notably less consistent
and can vary dramatically between individuals. We expect to see similar
trends in our dataset, but as stated previously, our observations
contain significantly more men than women. Finally, we wish to test
whether or not certain occupational industries have different effects on
sleep quality. It has been shown that managerial positions tend to have
the worst sleep quality among civilian sector workers while 24 jobs that
have rotating shifts tend to have the worst sleep quality [@luckhaupt_prevalence_2010]. 
Our dataset contains observations of mostly white-collar/service-based jobs, so 
we expect to find similar trends upon analysis.

Additionally, based on the findings of Sun et. al that obesity can
negatively impact sleep quality, we also expect that higher levels of
blood pressure will result in poorer sleep quality [@sun_poor_2016]. 
The American Sleep Foundation states that both short and long sleep duration 
increase risk of hypertension in adults [@liu_association_2022]. This draws 
interesting focus to the circular nature of sleep quality and hypertension. 
Those with hypertension also face a higher risk of insomnia, a disorder
affecting the ability to fall asleep, maintain sleep, and sleep without
continuous disturbance [@liu_association_2022]. Short sleep duration is also
positively associated with stress level [@noauthor_stress_nodate]. Survey findings
indicate that the average American sleeps 6.7 hours per night, which is
well below the minimum recommended duration seven to nine hours [@noauthor_stress_nodate]. 
Similar to the relationship between hypertension and sleep quality, as stress 
level increases, sleep duration decreases thereby causing poorer sleep quality. 
Poorer sleep quality then leads to higher stress, greater risk of hypertension, 
and shorter sleep duration.

@@

The dataset originally contained 373 observations for 13 variables
covering a broad spectrum of health and lifestyle variables associated
with sleep quality. The curators of the dataset did not provide
information on how or where this data was collected, so we solely relied
on the codebook for variable definitions. We have added this codebook to
the supplementary materials file for ease of access but you can also
find it on the Kaggle page for this dataset. There were no missing
values or erroneous variables in the dataset that we needed to remove.
The "Blood.Pressure" variable was transformed to reflect the individual
systolic and diastolic blood pressure measurements from each subject. We
then chose to add an additional variable, labeled 'cat_bp' to reflect
the categorical blood pressure status of a subject based on the
[American Heart Association
guidelines](https://www.heart.org/en/health-topics/high-blood-pressure) [@noauthor_understanding_nodate].
Additionally we created a variable named 'StepsGroup' to determine the
categorical activity level of subjects in the dataset based on these
guidelines provided by [10,000
steps](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps
/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,
day%205%20Highly%20active%20is%20more%20than%2012%2C500) [@noauthor_counting_nodate].
Finally, we created an additional categorical variable called
'PhysicalActivityGroup' to categorically represent the level of physical
activity reported by a subject. The levels are differentiated by 30
minute intervals of weekly physical activity. The outcome of interest is
noted as "Quality.of.Sleep" that we will reference as Sleep Quality or
Quality of Sleep throughout this report.
